Optimal updating factor in adaptive flat-distribution-sampling simulations

### 2017-05-10 doc/invt.tex
small edits to invt.tex

The above symmetry implies the existence of
a left eigenvector, $\pmb \rho$,
corresponding to the uniform right eigenvector
associated with Eq. \eqref{eq:w_sumj},
with eigenvalue $1$:
%
\begin{equation}
  \sum_{i = 1}^n \rho_i \, w_{ij}
  =
  \sum_{i = 1}^n \rho_j \, w_{ji}
  =
  \rho_j
  .
  \label{eq:w_balance}
\end{equation}
%
We will label this eigenvector as the first one,
and
%
\begin{equation}
  \phi_{1i} = \rho_i,
\label{eq:eigenmode1}
\end{equation}
%
with $\lambda_1 = 1$.
%
Clearly, Eq. \eqref{eq:eigenmode1}
satisfies the normalization condition
given by Eq. \eqref{eq:eig_orthonormal_rows}
for $k = l = 1$,
while the orthogonality for $k \ge 2$ demands
%
\begin{equation}
  \sum_{ i = 1 }^n \phi_{ki}
  =
  \sum_{ i = 1 }^n
    \frac{ \phi_{ki} \, \phi_{1i} }
         { \rho_i }
  =
  0
  .
\label{eq:ortho1}
\end{equation}

@@

We can generalize
the single-bin scheme to a class of
perfect (brick-wall) \emph{bandpass} updating schemes
that ignore short-wavelength eigenmodes,
which are typically local noise.
%
Mathematically, these updating schemes
allow us to assign zero eigenvalues
to modes that are likely absent from the PMF.
%
For example,
%if we can assume
by assuming
that the PMF is sufficiently smooth
such that the eigenmodes can be limited to $k \le K$,
%
we can construct an updating scheme with
$$
\lambda_1 = \cdots = \lambda_K = 1,
\mathrm{\; and \;}
\lambda_{K+1} = \cdots = \lambda_n = 0.
$$
%
Then,
we can change the upper limit of the sum in
Eq. \eqref{eq:error_asym_singlebin} to $K$:
%
%Then,
%Eq. \eqref{eq:error_singlebin} is changed to
%
\begin{equation}
  \Err_A(T)
  =
  \frac {       T     }
        { (T + t_0)^2 }
  \sum_{ k = 2 }^K
    \Gamma_k.
%\notag
\label{eq:error_asym_sinc}
\end{equation}
%
The optimal schedule of these schemes
remains to be the inverse-time schedule
given by Eq. \eqref{eq:alpha_invt1}.
%
Below, we construct these schemes for
the homogeneous case.
%
The definition of the cutoff, $K$, will be modified
for mathematical convenience.

\subsubsection{\label{sec:finlencorr}
Finite-length correction}

By including the residual error,
we determine the optimal $t_0$
that minimizes the total error.
%
According to Eq. \eqref{eq:half_alpha0}, we have
%
\begin{equation}
  t_0
  =
  \frac{        1      }
       { \alpha(t = 0) }
  =
  \frac{  2  }
       { a_0 }
  .
\label{eq:t0_sinc}
\end{equation}
%
For the single-bin updating scheme,
it recovers Eq. \eqref{eq:t0_singlebin}
under the assumption of Eq. \eqref{eq:y2_eql}.
%
\note{From Eq. \eqref{eq:y2_eql}, we have
$$E(0) = \frac{1}{2} a_0 \sum_{k=2}^n \Gamma_k = \frac{1}{2} \, a_0 \, \Gamma.$$
}
From Eqs. \eqref{eq:error_tot},
\eqref{eq:error_res1},
\eqref{eq:tk0},
and
\eqref{eq:error_asym_sinc},
we get
%
\begin{equation}
  \Err(T)
  =
  \frac{   1     }
       { T + t_0 }
  \sum_{ k = 2 }^K
    \Gamma_k
  .
\label{eq:error_sinc}
\end{equation}
%
In particular, for $K = n$, we have
\begin{equation}
  \sum_{k=2}^n \Gamma_k
  = \sum_{i=1}^n \rho_i \, G_i = \Gamma
  ,
  \notag
  %\label{eq:Gamma_G_sum}
\end{equation}
which can be shown by following the steps
similar to those leading to Eq. \eqref{eq:y2_sum},
and the above error recovers the single-bin result,
Eq. \eqref{eq:Emin_singlebin}.
%
%
%
\note{The derivation.
From Eq. \eqref{eq:eig_orthonormal_cols}, we have
$\sum_{i=1}^n \rho_i \langle x_i(t) \, x_i(0) \rangle
=\sum_{k=2}^n \rho_i \langle y_k(t) \, y_k(0) \rangle$,
so
\begin{align*}
\sum_{i=1}^n \rho_i G_i
&=\sum_{t=-\infty}^\infty\sum_{i=1}^n \rho_i \langle x_i(t) \, x_i(0) \rangle \\
&=\sum_{t=-\infty}^\infty\sum_{k=2}^n \rho_i \langle y_k(t) \, y_k(0) \rangle
=\sum_{k=2}^n \Gamma_k.
\end{align*}
}
%
This error is roughly proportional to the inverse simulation length,
as one would expect from an equilibrium FDS simulation.
%In Appendix \ref{sec:equilerr},
%we show that the above error tends
%to the error of the ideal equilibrium FDS simulation
%in the asymptotic limit.

@@

>>>>

The above symmetry implies the existence of
a left eigenvector, $\pmb \rho$,
corresponding to the uniform right eigenvector
associated with Eq. \eqref{eq:w_sumj},
with eigenvalue $1$:
%
\begin{equation}
  \sum_{i = 1}^n \rho_i \, w_{ij}
  =
  \sum_{i = 1}^n \rho_j \, w_{ji}
  =
  \rho_j
  .
  \label{eq:w_balance}
\end{equation}
%
We will label this eigenvector as the first one,
and
%
\begin{equation}
  \phi_{1i} = \rho_i,
\label{eq:eigenmode1}
\end{equation}
%
with $\lambda_1 = 1$.
%
Clearly, Eq. \eqref{eq:eigenmode1}
satisfies the normalization condition
given by Eq. \eqref{eq:eig_orthonormal_rows}
for $k = l = 1$,
while the orthogonality for $k \ge 2$ demands
%
\begin{equation}
  \sum_{ i = 1 }^n \phi_{ki}
  =
  \sum_{ i = 1 }^n
    \frac{ \phi_{ki} \, \phi_{1i} }
         { \rho_i }
  =
  0
  .
\label{eq:ortho1}
\end{equation}
%
We may write the updating matrix in terms of $\phi_{ki}$'s as
%
\begin{equation}
  w_{ij}
  =
  \frac{1}{\rho_i} \sum_{k=1}^n \lambda_k \, \phi_{ki} \, \phi_{kj}
  .
  \notag
\end{equation}

@@

We can generalize
the single-bin scheme to a class of
perfect (brick-wall) \emph{bandpass} updating schemes
that ignore short-wavelength eigenmodes,
which are typically local noise.
%
Mathematically, these updating schemes
allow us to assign zero eigenvalues
to modes that are likely absent from the PMF.
%
For example,
%if we can assume
by assuming
that the PMF is sufficiently smooth
such that the eigenmodes can be limited to $k \le K + 1$,
%
we can construct an updating scheme with
$$
\lambda_1 = \cdots = \lambda_{K+1} = 1,
\mathrm{\; and \;}
\lambda_{K+2} = \cdots = \lambda_n = 0.
$$
%
Then, the updating matrix is given by
%
$$w_{ij} = \sum_{k=1}^{K+1} \phi_{ki} \, \phi_{kj},$$
%
and Eq. \eqref{eq:error_asym_singlebin} is modified as
%
\begin{equation}
  \Err_A(T)
  =
  \frac {       T     }
        { (T + t_0)^2 }
  \sum_{ k = 2 }^{K+1}
    \Gamma_k.
%\notag
\label{eq:error_asym_sinc}
\end{equation}
%
The optimal schedule of these schemes
remains to be the inverse-time schedule
given by Eq. \eqref{eq:alpha_invt1},
%
with
%
\begin{equation}
  t_0
  =
  \frac{        1      }
       { \alpha(t = 0) }
  =
  \frac{  2  }
       { a_0 }
  ,
\label{eq:t0_sinc}
\end{equation}
%
where we have used Eq. \eqref{eq:half_alpha0}
in the second step.
%
\note{
For the single-bin updating scheme,
it recovers Eq. \eqref{eq:t0_singlebin}
under the assumption of Eq. \eqref{eq:y2_eql}.
%
From Eq. \eqref{eq:y2_eql}, we have
$$E(0) = \frac{1}{2} a_0 \sum_{k=2}^n \Gamma_k = \frac{1}{2} \, a_0 \, \Gamma.$$
}
%
From Eqs. \eqref{eq:error_tot},
\eqref{eq:error_res1},
\eqref{eq:tk0},
and
\eqref{eq:error_asym_sinc},
we get
%
\begin{equation}
  \Err(T)
  =
  \frac{   1     }
       { T + t_0 }
  \sum_{ k = 2 }^{ K + 1 }
    \Gamma_k
  .
\label{eq:error_sinc}
\end{equation}
%
This result recovers Eq. \eqref{eq:Emin_singlebin}
for the single-bin updating scheme:
if we set $K + 1 = n$,
%
\begin{equation}
  \sum_{k=2}^n \Gamma_k
  = \sum_{i=1}^n \rho_i \, G_i = \Gamma
  ,
  \notag
  %\label{eq:Gamma_G_sum}
\end{equation}
which can be shown by following the steps
similar to those leading to Eq. \eqref{eq:y2_sum}.
%
%
%
\note{The derivation.
From Eq. \eqref{eq:eig_orthonormal_cols}, we have
$\sum_{i=1}^n \rho_i \langle x_i(t) \, x_i(0) \rangle
=\sum_{k=2}^n \rho_i \langle y_k(t) \, y_k(0) \rangle$,
so
\begin{align*}
\sum_{i=1}^n \rho_i G_i
&=\sum_{t=-\infty}^\infty\sum_{i=1}^n \rho_i \langle x_i(t) \, x_i(0) \rangle \\
&=\sum_{t=-\infty}^\infty\sum_{k=2}^n \rho_i \langle y_k(t) \, y_k(0) \rangle
=\sum_{k=2}^n \Gamma_k.
\end{align*}
}
%
This error is roughly proportional to the inverse simulation length,
as one would expect from an equilibrium FDS simulation.
%In Appendix \ref{sec:equilerr},
%we show that the above error tends
%to the error of the ideal equilibrium FDS simulation
%in the asymptotic limit.

@@

\subsubsection{\label{sec:homo_bandpass}
Bandpass updating schemes}

For the bandpass updating schemes,
the eigenvalues are known and
we can use the inversion formula, Eq. \eqref{eq:mu_from_lambda}
to find the updating kernel.
%
Below we will slightly modify the definition of
the mode cutoff, $K$,
for mathematical convenience.
